---
title: "BUA_455_Week_12"
author: "Penelope Pooler Eisenbies"
date: "2022-11-15"
output: 
  html_document: 
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=TRUE}

# this line specifies options for default options for all R Chunks
knitr::opts_chunk$set(echo=T, 
                      highlight=T)

# suppress scientific notation
options(scipen=100,
        getSymbols.warning4.0 = FALSE)

# install helper package (pacman), if needed
if (!require("pacman")) install.packages("pacman", repos = "http://lib.stat.cmu.edu/R/CRAN/")

# install and load required packages
# pacman should be first package in parentheses and then list others
pacman::p_load(pacman,tidyverse, ggthemes, magrittr, knitr, lubridate, gridExtra, RColorBrewer, 
               flexdashboard, maps, usdata, countrycode, mapproj, shadowtext, plotly)

# verify packages (comment out in finished documents)
p_loaded()

```

---

### Updates

- HW 2 - Part 5 is due Friday, 11/18.  

  - For Question 9, you should document steps to update Panel 6.
  
  - Data for this panel were updated this week:
  
    - [Netflix Quarterly Revenue (requires SU login)](https://www-statista-com.libezproxy2.syr.edu/statistics/273883/netflixs-quarterly-revenue/)
    - [Netflix Quarterly Net Income (requires SU login)](https://www-statista-com.libezproxy2.syr.edu/statistics/273884/netflixs-quarterly-net-income/)

  - There is a grace period (over break), but you should be able to finish it before then.
  
---

### More Uppdates

- Quiz 2 grading is progressing

- Proposal Feedback will be posted this week.

  - Note: One group has asked to ammend their proposal and I agreed.  If your group is considering a switch, this would be the last week that it would be feasible.
  
- If you have data management questions, get in touch with me or a TA ASAP
  
  - Plan out tasks and how to accomplish them.  
  
  - We are here to help with tasks where you might be stymied, but don't wait until the last day.
  
- Presentations will be on 12/6 and 12/8.  

  - All students are required to attend and provide feedback.

---

### Plan for this Week

- More examples and skills to manage and curate geographic data

  - Today - World Data
  
  - Thursday - State data, Filtering to a Region, Publisng work

- More tips for good project management

- Posting HTML files for free using **[Rpubs](https://plotly.com/ggplot2/geom_polygon/)** 

  - Useful for presenting dashboard
  
---

### Importing and Joining World Datasets

#### World Data 

```{r world data prep}

world <- map_data("world") |> select(!subregion)                              # world geo info
intbxo <- read_csv("intl_bxo.csv", show_col_types = F, skip=6) |>             # import/tidy bxo
  select(1,6) |>
  rename("region" = "Area", "wknd_gross" = "Weekend Gross") |>
  filter(!is.na(wknd_gross)) |>
  mutate(wknd_gross = gsub("$", "", wknd_gross, fixed = T),
         wknd_gross = gsub(",", "", wknd_gross, fixed = T) |> as.numeric())

world_bxo_data <- left_join(intbxo, world) |>                                # join datasets
  filter(!is.na(wknd_gross))
world_bxo_data$continent = countrycode(sourcevar = world_bxo_data$region,    # retrieve continents
                                       origin = "country.name",
                                       destination = "continent")  
head(world_bxo_data, 3)

```
  
---

### Choropleth Country Plot w/ Labels

#### Example - Asia

- **Most** of the plot code that follows is review

  - There are a few new details:
    
    - `shadowtext` labels (see below)
    
    - modifying size of text elements (mentioned but not emphasized)

- **NOTES:** 

  - The R package `shadowtext` includes the command `geom_shadowtext`

  - `shadowtext` is useful for creating visible labels for all countries regardless of color

  - Deciding on units ($1000) and transformation (`log`) took some trial and error.
  
---
  
### Managing Data for Asia Chropleth Map

#### This R code creates the Asia Map dataset

```{r asia data for map}

asia_bxo_data <- world_bxo_data |>           # create asia box office dataset 
  filter(continent=="Asia") |>
  mutate(Gross = as.integer(wknd_gross), 
         wknd_gross = wknd_gross/1000) 

asia_nms <- asia_bxo_data |>                         # create dataset of country names 
  select(region, long, lat, group, continent) |>     # median lat and long used for position
  group_by(continent, region) |>
  summarize(nm_x=median(long, na.rm=T),
            nm_y=median(lat, na.rm=T)) |>
  filter(!is.na(nm_x) | !is.na(nm_y))

asia_bxo_data <- inner_join(asia_bxo_data, asia_nms) # merge datasets using an inner_join

```

---

### Asia Static Map with Log Transformation

```{r asia static map code}

(asia_bxo_map <- asia_bxo_data |>    # Creates the map that follows
   ggplot(aes(x=long, y=lat,
              group=group,
              fill=wknd_gross)) +
   geom_polygon() +
   theme_map() +
   coord_map("albers", lat0 = 39, lat1 = 45) +
   labs(fill= "Gross ($1K)",
        title="Weekend Gross ($ Thousands) in Asian Countries",
        subtitle="Weekend Ending 11/14/2022 - Data are Log-transformed",
        caption="Data Source: https://www.boxofficemojo.com") +
    
   scale_fill_continuous(type = "viridis",  trans="log",
                         breaks =c(1,10,100,1000,10000)) +
   geom_shadowtext(aes(x=nm_x, y=nm_y,label=region),
                   color="white",check_overlap = T,
                   show.legend = F, size=4) + 
                   
   theme(plot.title = element_text(size = 20),
         plot.subtitle = element_text(size = 15),
         plot.caption = element_text(size = 10),
         legend.text = element_text(size = 12),
         legend.title = element_text(size = 15),
         plot.background = element_rect(colour = "darkgrey", fill=NA, size=2)))

```


---

### Asia Interactive Map with Log Transformation

Map projection appears incorrect

```{r asia interactive map, warning=FALSE, message=FALSE}

asia_bxo_map_intr <- asia_bxo_data |>
   ggplot(aes(x=long, y=lat, group=group)) +
   geom_polygon(aes(fill=Gross)) +
   geom_line(aes(text=region, color=Gross), show.legend=F, alpha = 0.01) +  # used for interactive labels
   theme_map() +
   coord_map("albers", lat0 = 39, lat1 = 45) +
   labs(fill= "Gross", color="",
        title="Asian Countries - Weekend Gross - Weekend Ending 11/14/22"
        ) +
   scale_fill_continuous(type = "viridis",  trans="log",
                         breaks =c(1,10,100,1000,10000, 100000, 1000000),
                         labels=c("1", "10", "100", "1K", "10K", "100K", "1M")) +
   theme(plot.title = element_text(size = 15),
         legend.text = element_text(size = 10),
         legend.title = element_text(size = 10),
         legend.background = element_rect(colour = "lightgrey", fill=NA, size=0.5),
         panel.border = element_rect(colour = "lightgrey", fill=NA, size=2))

(asia_intr_map <- ggplotly(asia_bxo_map_intr, 
                           tooltip = c("text","color")))

```

---


### Lecture 19 In-class Exercise

- If you project includes data that can be mapped, this exercise will be helpful

- Details like text size, `shadowtext`, and data transformations can all greatly improve data clarity.

- You can choose to create a static map with shadow text or an interactive map

  - If you create static map, submit the .png file
  
  - If you create an interactive map, take screenshot with an interactive label showing

---

### Europe Box Office Map

- Copy and paste R code from previous chunks

- Convert it to create dataset for Europe map and one of the maps shown.

- This requires going through code line by line and making the following changes:

  - change `asia` to `euro`
  
  - change "Asia" to "Europe"
  
  - In static map, country labels in map may need to be slightly smaller (size=3)
  
---
  
    
### Europe Map Data

Creates data for Europe Map

```{r europe data for map}

euro_bxo_data <- world_bxo_data |>           # create europe box office dataset 
  filter(continent=="Europe" & region != "Russia") |>
  mutate(Gross = as.integer(wknd_gross), 
         wknd_gross = wknd_gross/1000) 

euro_nms <- euro_bxo_data |>                         # create dataset of country names 
  select(region, long, lat, group, continent) |>     # median lat and long used for position
  group_by(continent, region) |>
  summarize(nm_x=median(long, na.rm=T),
            nm_y=median(lat, na.rm=T)) |>
  filter(!is.na(nm_x) | !is.na(nm_y))

euro_bxo_data <- inner_join(euro_bxo_data, euro_nms) # merge datasets using an inner_join

```

---

### Europe Static Map with Log (LN) Transformation

Student will create this map or interactive one by adapting Asia Map code

```{r europe static map}

# copy, paste, and modify asia static map code in this chunk


```

---

### European Interactive Map with Log (LN) Transformation

Students will create this map or static one by adapting Asia Map code (projection is not correct).

```{r euro interactive map, warning=FALSE, message=FALSE}

# copy, paste, and modify asia interactive map code in this chunk

```


---

### Managing Projects 
  
- At this point, much of this should be review

- After break, I will spend a little time walking you through a consulting project

  - Project involved segmenting and rejoining data in multiple ways
  
  - Client asked for additional analyses months later

- Documentation is key

  - Take good notes and keep README file updated

- I use a Markdown files for everything, even work I don't present to client.
  
  - Ideal format for wiring notes between code
  
---

### What Goes where in a BUA 455 R Projects
  
#### **code_data_output folder** should have: 
  
  - All raw .csv files needed

  - ***NO data management should be done in Excel.***
  
  - Any .png or other graphics files needed
  
  - Dashboard .rmd file (R Markdown file)

  - Dashboard .html file (Dashboard presentation)
  
  - NOTE: As you progress in data management, code, data, and output may be in separate folders
    
    - requires specifying folder and file name
    
---

### What Goes where in a BUA 455 R Projects

#### **outer folder** should have

  - code_data_output folder (see above)

  - Complete and accurate README.txt file

  - .Rproj file that is functional 

- Other files can be stored in project **outer** file if needed

- If other files are stored in this outer folder they must be catalogued in the README file

---

### Publishing Work on Rpubs

#### **RPubs**

- If you want to publish your dashboard or any HTML file you create in R, you can do so for free.  

- R has a public online repository called [RPubs](https://rpubs.com/).  

- I am not requiring students to use it for their projects, but it is useful if you want post something online and provide the link to it.

- This is particularly useful if your dashboard is large and slow to knit

  - Being able to post the link instead of the .html file saves time

- As an in class exercise on Thursday, I will ask you each to create an account and publish your HW 5 - Part 1 dashboard html file 

  - This will be useful to experiment with
  
    - You can see how publishing changes the appearance of your panels and text.
  
    - In previous semesters, some students have saved their final dashboard online and linked it to their resumes and LinkedIn profiles.
    
---


### **Key Points from Week 12**
  
- **HW 2 - Part 2**

  - Due Friday, 11/18/22
    
- **More with Geographic Data**

   - Adding Shadow Text
   - Introduction to Interactive Maps
   - Projection Needs more tinkering

- **Project Management**

   - skills emphasized throughout course
   - managing course project this way is beneficial
   
- **Publishing Work on RPubs**

   - Useful for publishing and linking to work
    

**You may submit an 'Engagement Question or Comment' about Week 12 lectures until Thursday, 11/17, at midnight on Blackboard.**
